---
title: "How Camera Specs Influence Price Over Time"
author: "Team: Resolution Revolution"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

## Research Topic
This project investigates how camera specifications, especially megapixels, relate to camera price over time.

## Team Members
- Owen Sanders

## Data Description

I am using a camera dataset that includes specifications such as price, megapixels, release year, and weight.

**Dataset source:**  
https://raw.githubusercontent.com/prasertcbs/basic-dataset/master/camera_dataset.csv

### Variables of interest:
- Price 
- Max Resolution 
- Release Date 
- Brand 
- Weight

### Planned data cleaning steps:
- Rename variables for clarity
- Remove rows with missing values or zeros in price, weight, or resolution
- Extract brand name from model column

```{r}
url <- "https://raw.githubusercontent.com/prasertcbs/basic-dataset/master/camera_dataset.csv"
cameras_raw <- read_csv(url)

cameras_clean <- cameras_raw %>%
  rename(
    Release_Year = `Release date`,
    Max_Resolution = `Max resolution`,
    Weight_g = `Weight (inc. batteries)`
  ) %>%
  mutate(Brand = word(Model, 1)) %>%
  filter(Price > 0, Weight_g > 0, Max_Resolution > 0)

head(cameras_clean)
```

## Basic Exploratory Analysis

Will explore:
- Distribution of camera prices
- Relationship between megapixels and price
- Trends in megapixels over time

```{r}
ggplot(cameras_clean, aes(x = Price)) +
  geom_histogram(bins = 30, fill = "steelblue") +
  scale_x_continuous(labels = scales::dollar_format()) +
  labs(title = "Distribution of Camera Prices")
```

```{r}
ggplot(cameras_clean, aes(x = Max_Resolution, y = Price)) +
  geom_point(alpha = 0.5, color = "purple") +
  geom_smooth(method = "lm") +
  labs(title = "Megapixels vs Price")
```

## Project Research Questions

### 1. Does increasing megapixels actually lead to higher camera prices?
I will use regression analysis and scatterplots to evaluate whether higher resolution cameras cost more.

### 2. How has the relationship between megapixels and price changed over time?
I will explore whether newer cameras show weaker or stronger relationships between resolution and price by incorporating release year into some visualizations.

Planned analyses:

- Linear regression (Price ~ Megapixels)

- Color-coded scatterplots by year

- Faceted plots by time periods


# --- Answering RQ1: Linear Regression ---
# Does increasing megapixels actually lead to higher camera prices?
```{r}
model_1 <- lm(Price ~ Max_Resolution, data = cameras_clean)
summary(model_1) 
```
# Look at the 'Estimate' for Max_Resolution to see how much price goes up per megapixel!

# --- Answering RQ2: Trends over time ---
# How has the relationship changed over time? Color-coded scatterplot.
```{r}
ggplot(cameras_clean, aes(x = Max_Resolution, y = Price, color = Release_Year)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", color = "black", linetype = "dashed") +
  scale_color_viridis_c() + # Makes the years look nice and distinct
  labs(title = "Impact of Megapixels on Price Over Time",
       x = "Max Resolution (Megapixels)",
       y = "Price ($)",
       color = "Release Year") +
  theme_minimal()
```